---
title: "GLMM – Verbal Fluency & ROI Activation"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Import modules

```{r}
library(lme4)         # mixed model package
library(lmerTest)     # p-values for mixed models
library(ggplot2)      # graphics
library(emmeans)      # post-hoc tests
library(dplyr)
library(tidyr)
library(readr)
library(openxlsx)
library(broom.mixed)  # tidy() for mixed models

script_dir <- tryCatch(
  dirname(rstudioapi::getSourceEditorContext()$path),
  error = function(e) getwd()
)
out_dir <- file.path(script_dir, "results")
if (!dir.exists(out_dir)) dir.create(out_dir)
```

# Speech–ROI GLMM Analysis

One GLMM per ROI with all speech features as predictors: `mean_contrast ~ word_count_z + speech_ratio_z + pause_rate_z + pause_durations_mean_z + type_token_ratio_z + brunets_index_z + group + age + sex + hemisphere + (1 | subject)`

## Read in dataset

```{r}
# ---- ROI data ----
roi_path <- file.path(dirname(script_dir), "new_mri", "ROI", "results",
                      "roi_condition_values_long.csv")
roi_raw <- read_csv(roi_path, show_col_types = FALSE)
roi_raw <- roi_raw[!is.na(roi_raw$mean_contrast), ]
roi_raw$region <- sub("^[LR]_", "", roi_raw$roi)

# Average across 4 conditions → 1 value per subject × region × hemisphere
roi_avg <- roi_raw %>%
  group_by(subject, group, region, hemisphere) %>%
  summarise(mean_contrast = mean(mean_contrast), .groups = "drop")

# ---- Speech data ----
speech_raw <- read.csv(file.path(script_dir, "combined_speech_demographic_info.csv"),
                       sep = ";", stringsAsFactors = FALSE)
speech_raw$subject <- tolower(speech_raw$bids_number)

fluency_features <- c("word_count", "speech_ratio", "pause_rate",
                       "pause_durations_mean", "type_token_ratio", "brunets_index")

speech_df <- data.frame(subject = speech_raw$subject)
for (feat in fluency_features) {
  pos_vals <- as.numeric(speech_raw[[paste0(feat, "_pos")]])
  neg_vals <- as.numeric(speech_raw[[paste0(feat, "_neg")]])
  speech_df[[feat]] <- rowMeans(cbind(pos_vals, neg_vals), na.rm = TRUE)
}

# ---- Demographics: age & sex ----
speech_df$age <- as.numeric(speech_raw[["Alter"]])
speech_df$sex <- factor(speech_raw[["Gender_f1_m2"]], levels = c(1, 2), labels = c("f", "m"))
```

```{r}
# ---- Merge & factor coding ----
merged <- inner_join(roi_avg, speech_df, by = "subject")
merged$group      <- factor(merged$group, levels = c("HC", "MDD"))
merged$hemisphere <- factor(merged$hemisphere, levels = c("L", "R"))
merged$subject    <- factor(merged$subject)

regions    <- sort(unique(merged$region))
n_models   <- length(regions)   # = 3

# Z-score all speech features
for (feat in fluency_features) {
  merged[[paste0(feat, "_z")]] <- scale(merged[[feat]])[, 1]
}
feat_z <- paste0(fluency_features, "_z")
```

```{r}
# Count unique participants in each group
merged %>%
  group_by(group) %>%
  summarise(Num_Participants = n_distinct(subject))
```

## Descriptive: Speech Features by Group

```{r}
speech_subj <- merged %>%
  select(subject, group, all_of(fluency_features)) %>%
  distinct()

desc <- speech_subj %>%
  pivot_longer(all_of(fluency_features), names_to = "feature", values_to = "value") %>%
  group_by(group, feature) %>%
  summarise(M = mean(value, na.rm = TRUE), SD = sd(value, na.rm = TRUE), .groups = "drop")

print(desc)
write.csv(desc, file.path(out_dir, "speech_descriptives_by_group.csv"), row.names = FALSE)
```

## Estimate GLMMs

### ACC

```{r}
d_acc <- merged[merged$region == "ACC", ]

model_acc <- lmer(mean_contrast ~ word_count_z + speech_ratio_z + pause_rate_z +
                    pause_durations_mean_z + type_token_ratio_z + brunets_index_z +
                    group + age + sex + (1 | subject),
                  data = d_acc)

anova(model_acc)
```

```{r}
coefs_acc <- tidy(model_acc, effects = "fixed", conf.int = TRUE, conf.method = "Wald")
coefs_acc$p_bonf <- pmin(coefs_acc$p.value * 3, 1)   # Bonferroni × 3 models
coefs_acc$sig <- ifelse(coefs_acc$p_bonf < .001, "***",
                 ifelse(coefs_acc$p_bonf < .01, "**",
                 ifelse(coefs_acc$p_bonf < .05, "*", "")))
coefs_acc$Region <- "ACC"
coefs_acc
```

```{r}
eff_acc <- emmeans(model_acc, pairwise ~ group, adjust = "bonferroni")
summary(eff_acc$emmeans)
summary(eff_acc$contrasts)
```

### Amy

```{r}
d_amy <- merged[merged$region == "Amy", ]

model_amy <- lmer(mean_contrast ~ word_count_z + speech_ratio_z + pause_rate_z +
                    pause_durations_mean_z + type_token_ratio_z + brunets_index_z +
                    group + age + sex + (1 | subject),
                  data = d_amy)

anova(model_amy)
# test significant speech features again with group interaction
model_amy_2 <- lmer(mean_contrast ~ speech_ratio_z*group + pause_rate_z*group +
                    pause_durations_mean_z*group + group + age + sex + (1 | subject),
                  data = d_amy)
anova(model_amy_2)
```

```{r}
coefs_amy <- tidy(model_amy, effects = "fixed", conf.int = TRUE, conf.method = "Wald")
coefs_amy$p_bonf <- pmin(coefs_amy$p.value * 3, 1)
coefs_amy$sig <- ifelse(coefs_amy$p_bonf < .001, "***",
                 ifelse(coefs_amy$p_bonf < .01, "**",
                 ifelse(coefs_amy$p_bonf < .05, "*", "")))
coefs_amy$Region <- "Amy"
coefs_amy
```

```{r}
eff_amy <- emmeans(model_amy, pairwise ~ group, adjust = "bonferroni")
summary(eff_amy$emmeans)
summary(eff_amy$contrasts)
```

### DLPFC

```{r}
d_dlpfc <- merged[merged$region == "DLPFC", ]

model_dlpfc <- lmer(mean_contrast ~ word_count_z + speech_ratio_z + pause_rate_z +
                      pause_durations_mean_z + type_token_ratio_z + brunets_index_z +
                      group + age + sex + (1 | subject),
                    data = d_dlpfc)

anova(model_dlpfc)
```

```{r}
coefs_dlpfc <- tidy(model_dlpfc, effects = "fixed", conf.int = TRUE, conf.method = "Wald")
coefs_dlpfc$p_bonf <- pmin(coefs_dlpfc$p.value * 3, 1)
coefs_dlpfc$sig <- ifelse(coefs_dlpfc$p_bonf < .001, "***",
                   ifelse(coefs_dlpfc$p_bonf < .01, "**",
                   ifelse(coefs_dlpfc$p_bonf < .05, "*", "")))
coefs_dlpfc$Region <- "DLPFC"
coefs_dlpfc
```

```{r}
eff_dlpfc <- emmeans(model_dlpfc, pairwise ~ group, adjust = "bonferroni")
summary(eff_dlpfc$emmeans)
summary(eff_dlpfc$contrasts)
```

## Combine Results

```{r}
coef_all <- bind_rows(coefs_acc, coefs_amy, coefs_dlpfc)

sig <- coef_all[coef_all$p_bonf < 0.05, ]
if (nrow(sig) > 0) {
  print(sig)
} else {
  cat("No significant fixed effects after Bonferroni correction.\n")
}
```

```{r}
# ANOVA tables (Type III, Kenward-Roger)
anova_acc   <- as.data.frame(anova(model_acc, type = 3, ddf = "Kenward-Roger"))
anova_acc$Term   <- rownames(anova_acc); anova_acc$Region   <- "ACC"

anova_amy   <- as.data.frame(anova(model_amy, type = 3, ddf = "Kenward-Roger"))
anova_amy$Term   <- rownames(anova_amy); anova_amy$Region   <- "Amy"

anova_amy2  <- as.data.frame(anova(model_amy_2, type = 3, ddf = "Kenward-Roger"))
anova_amy2$Term  <- rownames(anova_amy2); anova_amy2$Region  <- "Amy (interaction model)"

anova_dlpfc <- as.data.frame(anova(model_dlpfc, type = 3, ddf = "Kenward-Roger"))
anova_dlpfc$Term <- rownames(anova_dlpfc); anova_dlpfc$Region <- "DLPFC"

anova_all <- bind_rows(anova_acc, anova_amy, anova_amy2, anova_dlpfc)
anova_all
```

```{r}
fit_all <- data.frame(
  Region = c("ACC", "Amy", "DLPFC"),
  AIC    = c(AIC(model_acc), AIC(model_amy), AIC(model_dlpfc)),
  BIC    = c(BIC(model_acc), BIC(model_amy), BIC(model_dlpfc)),
  logLik = c(as.numeric(logLik(model_acc)), as.numeric(logLik(model_amy)), as.numeric(logLik(model_dlpfc)))
)
fit_all
```

```{r}
emmeans_acc   <- as.data.frame(eff_acc$emmeans);   emmeans_acc$Region   <- "ACC"
emmeans_amy   <- as.data.frame(eff_amy$emmeans);   emmeans_amy$Region   <- "Amy"
emmeans_dlpfc <- as.data.frame(eff_dlpfc$emmeans); emmeans_dlpfc$Region <- "DLPFC"
emmeans_all   <- bind_rows(emmeans_acc, emmeans_amy, emmeans_dlpfc)
emmeans_all

contrasts_acc   <- as.data.frame(eff_acc$contrasts);   contrasts_acc$Region   <- "ACC"
contrasts_amy   <- as.data.frame(eff_amy$contrasts);   contrasts_amy$Region   <- "Amy"
contrasts_dlpfc <- as.data.frame(eff_dlpfc$contrasts); contrasts_dlpfc$Region <- "DLPFC"
contrasts_all   <- bind_rows(contrasts_acc, contrasts_amy, contrasts_dlpfc)
contrasts_all
```

## Plot ACC

```{r}
plot_df <- merged %>%
  group_by(subject, group, region) %>%
  summarise(mean_contrast = mean(mean_contrast), .groups = "drop") %>%
  left_join(speech_subj, by = c("subject", "group"))
```

```{r}
d <- plot_df[plot_df$region == "ACC", ]
for (feat in fluency_features) {
  p <- ggplot(d, aes(x = .data[[feat]], y = mean_contrast, colour = group)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_smooth(method = "lm", se = TRUE, alpha = 0.15) +
    scale_colour_manual(values = c("HC" = "#56B4E9", "MDD" = "#CC79A7")) +
    labs(title = sprintf("ACC – %s", feat),
         x = feat, y = "Mean contrast", colour = "Group") +
    theme_minimal(base_size = 13)
  print(p)
}
```

## Plot Amy

```{r}
d <- plot_df[plot_df$region == "Amy", ]
for (feat in fluency_features) {
  p <- ggplot(d, aes(x = .data[[feat]], y = mean_contrast, colour = group)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_smooth(method = "lm", se = TRUE, alpha = 0.15) +
    scale_colour_manual(values = c("HC" = "#56B4E9", "MDD" = "#CC79A7")) +
    labs(title = sprintf("Amy – %s", feat),
         x = feat, y = "Mean contrast", colour = "Group") +
    theme_minimal(base_size = 13)
  print(p)
}
```

## Export to Excel

```{r}
# Save coefficients as CSV
write.csv(coef_all, file.path(out_dir, "speech_roi_glmm_coefficients.csv"), row.names = FALSE)

# Save all results to Excel
wb <- createWorkbook()

addWorksheet(wb, "Fixed Effects")
writeData(wb, "Fixed Effects", coef_all)

addWorksheet(wb, "Significant Effects")
if (nrow(sig) > 0) {
  writeData(wb, "Significant Effects", sig)
} else {
  writeData(wb, "Significant Effects",
            data.frame(Note = "No significant effects after Bonferroni correction"))
}

addWorksheet(wb, "Emmeans")
writeData(wb, "Emmeans", emmeans_all)

addWorksheet(wb, "Contrasts")
writeData(wb, "Contrasts", contrasts_all)

addWorksheet(wb, "ANOVA")
writeData(wb, "ANOVA", anova_all)

addWorksheet(wb, "Model Fit")
writeData(wb, "Model Fit", fit_all)

addWorksheet(wb, "Descriptives by Group")
writeData(wb, "Descriptives by Group", desc)

xlsx_path <- file.path(out_dir, "speech_roi_glmm_results.xlsx")
saveWorkbook(wb, xlsx_path, overwrite = TRUE)

cat(sprintf("Saved to:\n  %s\n", xlsx_path))
```

```{r}
# ANOVA tables (Type III, Kenward-Roger)
anova_acc   <- as.data.frame(anova(model_acc, type = 3, ddf = "Kenward-Roger"))
anova_acc$Term   <- rownames(anova_acc); anova_acc$Region   <- "ACC"

anova_amy   <- as.data.frame(anova(model_amy, type = 3, ddf = "Kenward-Roger"))
anova_amy$Term   <- rownames(anova_amy); anova_amy$Region   <- "Amy"

anova_amy2  <- as.data.frame(anova(model_amy_2, type = 3, ddf = "Kenward-Roger"))
anova_amy2$Term  <- rownames(anova_amy2); anova_amy2$Region  <- "Amy (interaction model)"

anova_dlpfc <- as.data.frame(anova(model_dlpfc, type = 3, ddf = "Kenward-Roger"))
anova_dlpfc$Term <- rownames(anova_dlpfc); anova_dlpfc$Region <- "DLPFC"

anova_all <- bind_rows(anova_acc, anova_amy, anova_amy2, anova_dlpfc)
anova_all
```

```{r}
# Save coefficients as CSV
write.csv(coef_all, file.path(out_dir, "speech_roi_glmm_coefficients.csv"), row.names = FALSE)

# Save all results to Excel
wb <- createWorkbook()

addWorksheet(wb, "Fixed Effects")
writeData(wb, "Fixed Effects", coef_all)

addWorksheet(wb, "Significant Effects")
if (nrow(sig) > 0) {
  writeData(wb, "Significant Effects", sig)
} else {
  writeData(wb, "Significant Effects",
            data.frame(Note = "No significant effects after Bonferroni correction"))
}

addWorksheet(wb, "Emmeans")
writeData(wb, "Emmeans", emmeans_all)

addWorksheet(wb, "Contrasts")
writeData(wb, "Contrasts", contrasts_all)

addWorksheet(wb, "ANOVA")
writeData(wb, "ANOVA", anova_all)

addWorksheet(wb, "Model Fit")
writeData(wb, "Model Fit", fit_all)

addWorksheet(wb, "Descriptives by Group")
writeData(wb, "Descriptives by Group", desc)

xlsx_path <- file.path(out_dir, "speech_roi_glmm_results.xlsx")
saveWorkbook(wb, xlsx_path, overwrite = TRUE)

cat(sprintf("Saved to:\n  %s\n", xlsx_path))
```
