---
title: "BDI Mediation on Speech and Brain Correlation"
output: html_notebook
---

```{r}
library(CCA) #perform canonical correlation
library(psych)
library(lavaan)
library(R.matlab) #read in matlab variable
library(CCP) #test which correlations are significant
library(ggplot2)# Load ggplot2 for advanced visualization
library(afex) 
library(cowplot)
library(prism)
library(dplyr) # to use %>% 
#install.packages("tidyverse")
library(tidyverse) # to transform data 
library(reshape2)
library(mediation) # to test mediation effect of BDI
```

```{r}
df <- read.csv("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/speech_and_ROI.csv")
X <- df[, c(
  "pos_frequency_related", "pos_energy_related", "pos_spectral_balance", 
  "pos_temporal", "pos_lexical_richness", "pos_sentiment", 
  "pos_word_types", "pos_syntactic_complexity", 
  "neg_frequency_related", "neg_energy_related", "neg_spectral_balance", 
  "neg_temporal", "neg_lexical_richness", "neg_sentiment", 
  "neg_word_types", "neg_syntactic_complexity"
)]

Y <- df[, c("AP_aInsula_BA13", "AP_Amygdala","AP_aSTG_BA22","AP_aSTG_BA22_2",
"AP_Caudate_nucleus","AP_Cerebellum","AP_IFG_pOp_BA44","AP_IFG_pOp_BA44_2",
"AP_IFG_pOrb_BA47","AP_IFG_pOrb_BA47_2","AP_IFG_pOrb_BA47_3","AP_IFG_pTri_BA45",
"AP_MFG_BA9","AP_MFG_BA10","AP_MTG_BA21","AP_Parahippocampal_gyrus_BA28",
"AP_pSTG_BA22","AP_pSTG_BA22_2","AP_pSTG_BA22_3","AP_Putamen","AP_SMA_BA6",
"AP_SMA_BA6_3","AP_SMG_BA40","AP_SMG_BA40_7","AP_SMG_BA40_7_2","AP_Subcallosal_gyrus_BA34",
"AP_Thalamus","LP_aInsula_BA13","LP_aInsula_BA13_2","LP_Caudate_nucleus","LP_Cerebellum",
"LP_Cerebellum_2","LP_Cerebellum_3","LP_Claustrum","LP_Cuneus_BA17",
"LP_Heschls_gyrus_BA41","LP_IFG_pOp_BA44","LP_IFG_pOp_BA44_2","LP_IFG_pOrb_BA47",
"LP_Insula_BA13","LP_MFG_BA9","LP_MFG_BA9_2","LP_pSTG_BA22","LP_pSTG_BA22_2","LP_SMA_BA6", "LP_SMA_BA6_2","LP_SMG_BA40","LP_SMG_BA40_2","LP_SMG_BA40_7"
)]
```

```{r}
X_scaled <- scale(X) #speech
Y_scaled <- scale(Y) #brain
cca_result <- cc(X_scaled, Y_scaled)
```

## Test for Mediation of Correlations thru BDI

Data preparation

```{r}
# Compute canonical scores
X_scores <- X_scaled %*% cca_result$xcoef
Y_scores <- Y_scaled %*% cca_result$ycoef

# Convert to data frames for easier manipulation
X_scores_df <- as.data.frame(X_scores)
Y_scores_df <- as.data.frame(Y_scores)

# Name the canonical scores for clarity (e.g., Canonical1, Canonical2, ...)
colnames(X_scores_df) <- paste0("Canonical_Speech", 1:ncol(X_scores_df))
colnames(Y_scores_df) <- paste0("Canonical_Brain", 1:ncol(Y_scores_df))

# Combine the canonical scores with BDI scores
mediation_data <- cbind(X_scores_df, Y_scores_df, BDI = df$BDI)
```

Mediation Models

This step is modeling how the **BDI** (mediator) is influenced by the **Canonical_Speech1**.
The idea is to understand if and how the speech features, as captured by the first canonical variate, impact the BDI scores.
If **Canonical_Speech1** significantly predicts **BDI**, it indicates that the speech features might influence the depression scores.

```{r}
# Mediator model: BDI as a function of Canonical_Speech1
model_m <- lm(BDI ~ Canonical_Speech1, data = mediation_data)
```

```{r}
# Outcome model: Canonical_Brain1 as a function of Canonical_Speech1 and BDI
model_y <- lm(Canonical_Brain1 ~ Canonical_Speech1 + BDI, data = mediation_data)
```

This model is designed to assess the relationship between the brain activity (captured by **Canonical_Brain1**) and the speech features (through **Canonical_Speech1**), while also including **BDI** as a potential mediator.

-   **Direct Effect**: The effect of **Canonical_Speech1** on **Canonical_Brain1**, controlling for **BDI**.

-   **Indirect Effect**: The effect of **Canonical_Speech1** on **Canonical_Brain1** that is mediated by **BDI**.

### Now test for existing BDI mediation for the first 4 canonicals

```{r}
print(colnames(mediation_data))

```

```{r}
if (!(treat %in% colnames(mediation_data))) {
    stop(paste("The treatment column", treat, "does not exist in the data."))
}

```

```{r}
# Initialize an empty list to store results
mediation_results_list <- list()

# Perform mediation analysis for the first 4 canonical variates
for (i in 1:11) {
    canonical_speech <- paste0("Canonical_Speech", i)
    canonical_brain <- paste0("Canonical_Brain", i)

    # Fit the mediator model
    model_m <- lm(as.formula(paste("BDI ~", canonical_speech)), data = mediation_data)
    
    # Fit the outcome model
    model_y <- lm(as.formula(paste(canonical_brain, "~", canonical_speech, "+ BDI")), data = mediation_data)
    
    # Perform mediation analysis
    med_fit <- mediate(model_m, model_y, treat = canonical_speech, mediator = "BDI", boot = TRUE, sims = 1000)

    # Store results
    mediation_results_list[[i]] <- data.frame(
        Canonical_Variate = paste0("Canonical_", i),
        ACME = med_fit$d0,
        ADE = med_fit$z0,
        Total_Effect = med_fit$tau.coef,
        Proportion_Mediated = med_fit$n0,
        CI_Lower = med_fit$d0.ci[1],
        CI_Upper = med_fit$d0.ci[2],
        p_value = med_fit$d0.p
    )
}

# Combine all mediation results into one data frame
mediation_results_df <- do.call(rbind, mediation_results_list)

# Write the results to a CSV file
write.csv(mediation_results_df, "mediation_results.csv", row.names = FALSE)

# Inform the user that the process is complete
cat("Mediation analysis results have been saved to 'mediation_results.csv'.\n")
```

-   **ACME (Average Causal Mediation Effect)**: This tells you how much of the effect of `Canonical_Speech1` on `Canonical_Brain1` is mediated by the BDI scores.

-   **ADE (Average Direct Effect)**: This measures the effect of `Canonical_Speech1` on `Canonical_Brain1` that is not mediated by BDI.

-   **Total Effect**: This is the sum of the ACME and ADE, representing the overall effect of `Canonical_Speech1` on `Canonical_Brain1`.

-   **Proportion Mediated**: This gives the proportion of the total effect that is mediated by BDI.

<!-- -->

-   **Significant Mediation**: The mediation effect (ACME) is small but significant, suggesting that BDI does mediate part of the relationship between the speech and brain canonical variates, albeit a small portion.

-   **Strong Direct Effect**: The direct effect (ADE) is very strong and highly significant, indicating that the majority of the relationship between the second canonical speech and brain variates is direct, independent of BDI scores.

-   **Implications**: While BDI has a small mediation effect, the primary driver of the association between the speech and brain canonical variates in this dimension seems to be direct rather than mediated by BDI.
