---
title: "speech_brain_correlation"
output: html_notebook
---

## Install required packages

```{r}
#install.packages("rainbow")
#install.packages("fds")
#install.packages("CCA")
#install.packages("tmvnsim")
#install.packages("psych")
#install.packages("lavaan", type = "binary")
#install.packages("R.matlab")
#install.packages("dplyr")  # For data manipulation
#install.packages("CCP")
```

```{r}
# Load necessary libraries
library(CCA) #perform canonical correlation
library(psych)
library(lavaan)
library(R.matlab) #read in matlab variable
library(CCP) #test which correlations are significant
```

```{r}
library(ggplot2)# Load ggplot2 for advanced visualization
library(afex) 
library(cowplot)
library(prism)
library(dplyr) # to use %>% 
#install.packages("tidyverse")
library(tidyverse) # to transform data 
library(reshape2)
library(mediation) # to test mediation effect of BDI
```

## Read in and Transform fMRI Data

```{r}
fmri_data <- readMat("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/extracted_roi_data.mat")
```

```{r}
## Accessing the first element of the main list
#first_element <- fmri_data[[1]]
## Accessing a nested element
#nested_element <- fmri_data$extracted.data[[1]][[2]][[3]]
```

```{r}
extracted_data <- fmri_data$extracted.data

# Initialize an empty list to store rows
data_list <- list()

# Loop over each ROI
for (roi_index in seq_along(extracted_data)) {
  roi_data <- extracted_data[[roi_index]]
  
  # Loop over each task condition within the current ROI
  for (condition_index in seq_along(roi_data)) {
    condition_data <- roi_data[[condition_index]]
    
    # Loop over each participant within the current task condition
    for (participant_index in seq_along(condition_data)) {
      brain_value <- condition_data[[participant_index]][1,1]  # Extract the numeric value
      
      # Store the data as a row in the list
      data_list[[length(data_list) + 1]] <- data.frame(
        ROI = roi_index,
        TaskCondition = condition_index,
        Participant = participant_index,
        BrainValue = brain_value
      )
    }
  }
}

# Combine all rows into a single data frame
combined_df <- do.call(rbind, data_list)

# Convert to a data frame
combined_df <- as.data.frame(combined_df)

# Convert columns to appropriate types
combined_df$ROI <- as.factor(combined_df$ROI)
combined_df$TaskCondition <- as.factor(combined_df$TaskCondition)
combined_df$Participant <- as.factor(combined_df$Participant)
combined_df$BrainValue <- as.numeric(combined_df$BrainValue)

# View the resulting data frame
head(combined_df)
```

```{r}
rm(extracted_data,fmri_data,condition_data,data_list,roi_data)
```

```{r}
write.csv(combined_df, "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/extracted_data.csv", row.names = FALSE)
```

```{r}
# List of ROI names based on the provided information
roi_names <- c("AP_aInsula_BA13", "AP_Amygdala","AP_aSTG_BA22","AP_aSTG_BA22_2",
"AP_Caudate_nucleus","AP_Cerebellum","AP_IFG_pOp_BA44","AP_IFG_pOp_BA44_2",
"AP_IFG_pOrb_BA47","AP_IFG_pOrb_BA47_2","AP_IFG_pOrb_BA47_3","AP_IFG_pTri_BA45",
"AP_MFG_BA9","AP_MFG_BA10","AP_MTG_BA21","AP_Parahippocampal_gyrus_BA28",
"AP_pSTG_BA22","AP_pSTG_BA22_2","AP_pSTG_BA22_3","AP_Putamen","AP_SMA_BA6",
"AP_SMA_BA6_3","AP_SMG_BA40","AP_SMG_BA40_7","AP_SMG_BA40_7_2","AP_Subcallosal_gyrus_BA34",
"AP_Thalamus","LP_aInsula_BA13","LP_aInsula_BA13_2","LP_Caudate_nucleus","LP_Cerebellum",
"LP_Cerebellum_2","LP_Cerebellum_3","LP_Claustrum","LP_Cuneus_BA17",
"LP_Heschls_gyrus_BA41","LP_IFG_pOp_BA44","LP_IFG_pOp_BA44_2","LP_IFG_pOrb_BA47",
"LP_Insula_BA13","LP_MFG_BA9","LP_MFG_BA9_2","LP_pSTG_BA22","LP_pSTG_BA22_2","LP_SMA_BA6", "LP_SMA_BA6_2","LP_SMG_BA40","LP_SMG_BA40_2","LP_SMG_BA40_7")
```

```{r}
participant_names <- c(
  "sub004", "sub006", "sub010", "sub011", "sub014", "sub015", "sub016", 
  "sub017", "sub018", "sub019", "sub022", "sub024", "sub025", "sub027", 
  "sub028", "sub029", "sub030", "sub031", "sub032", "sub033", "sub034", 
  "sub041", "sub043", "sub045", "sub046", "sub047", "sub050", "sub051", 
  "sub052", "sub053", "sub054", "sub056", "sub057", "sub059", "sub062", 
  "sub068", "sub069", "sub070", "sub071", "sub073", "sub074", "sub079", 
  "sub080", "sub083", "sub085", "sub086", "sub089", "sub090", "sub091", 
  "sub093", "sub096", "sub101", "sub103", "sub105", "sub119", "sub123", 
  "sub125", "sub126", "sub127", "sub007", "sub008", "sub009", "sub012", 
  "sub020", "sub035", "sub036", "sub037", "sub038", "sub039", "sub042", 
  "sub044", "sub048", "sub049", "sub061", "sub064", "sub065", "sub066", 
  "sub072", "sub075", "sub076", "sub077", "sub081", "sub082", "sub084", 
  "sub087", "sub092", "sub094", "sub095", "sub097", "sub098", "sub100", 
  "sub102", "sub104", "sub106", "sub107", "sub108", "sub109", "sub110", 
  "sub111", "sub113", "sub114", "sub115", "sub116", "sub117", "sub118", 
  "sub121", "sub122", "sub124", "sub128", "sub129", "sub130", "sub131"
)
```

```{r}
condition_names <- c("happy_happy", "happy_sad", "sad_happy", "sad_sad")
```

```{r}
# Load the CSV file you provided
combined_df_new <- combined_df
# Replace the ROI indices with the corresponding ROI names
combined_df_new$ROI <- roi_names[combined_df_new$ROI]
# Replace the task condition indices with the corresponding condition names
combined_df_new$TaskCondition <- condition_names[combined_df_new$TaskCondition]
# Replace the participant indices with the corresponding participant names
combined_df_new$Participant <- participant_names[combined_df_new$Participant]
# View the updated data frame to check if the mapping was successful
head(combined_df_new)
rm(combined_df)
```

## Descriptive (fMRI Data)

```{r}
# Summarize the data to get an overview
combined_df <- combined_df_new
summary(combined_df)

# Calculate mean brain activity by ROI and condition
mean_activity <- aggregate(BrainValue ~ ROI + TaskCondition, data = combined_df, FUN = mean)
print(mean_activity)

# Calculate standard deviations, medians, etc.
sd_activity <- aggregate(BrainValue ~ ROI + TaskCondition, data = combined_df, FUN = sd)
print(sd_activity)
rm(combined_df_new, mean_activity, sd_activity)
```

```{r, fig.width=15, fig.height=25}
# Plot mean brain activity for each ROI by condition
brain <- ggplot(combined_df, aes(x = TaskCondition, y = BrainValue, fill = TaskCondition)) +
  geom_boxplot() +
  facet_wrap(~ROI, scales = "free_y", nrow = 7) +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 16),
        axis.text= element_text(size = 16))+
  labs(title = "Brain Activity by ROI and Condition", y = "Brain Activity", x = "Conditions")
#ggsave("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/plots/brain_values.tiff", brain, width = 30, height = 18)
rm(brain)
```

## Read in Speech Data

```{r}
speech_data <- read.csv("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/speech/speech_categories.csv", sep=",")
head(speech_data)
```

```{r}
speech_data <- speech_data %>% 
  rename(id = bids_number) %>% 
  mutate(
    id = gsub("-", "", id),
    id = gsub("S", "s", id)
  )
head(speech_data)
```

```{r}
fmri_data <- combined_df %>% 
  rename(id = Participant) %>% 
  
  # Filter rows where TaskCondition ends with "happy" or "sad"
  filter(grepl("_happy$", TaskCondition) | grepl("_sad$", TaskCondition)) %>%
  
  # Create a new column 'NewCondition' to replace "_happy" and "_sad" with "happy" and "sad"
  mutate(NewCondition = ifelse(grepl("_happy$", TaskCondition), "happy", "sad")) %>%
  
  # Group by ROI and NewCondition
  group_by(ROI, NewCondition,id) %>%
  
  # Summarize to calculate the mean BrainValue for each group
  summarize(mean_BrainValue = mean(BrainValue, na.rm = TRUE))

head(fmri_data)
#rm(combined_df)
```

# Canonical Correlation Analysis (CCA)

```{r}
wide_fmri_data <- fmri_data %>%
  pivot_wider(
    names_from = c(ROI), # Columns that will become new column names
    values_from = mean_BrainValue            # Column that contains the values to spread
  )
head(wide_fmri_data)
# Find common participants
common_participants <- intersect(wide_fmri_data$id, speech_data$id)

# Filter both datasets to include only common participants
wide_fmri_data_filtered <- wide_fmri_data[wide_fmri_data$id %in% common_participants, ]
speech_data_filtered <- speech_data[speech_data$id %in% common_participants, ]

# Check column names in both datasets
#colnames(wide_fmri_data_filtered)
#colnames(speech_data_filtered)
rm(speech_data,wide_fmri_data)
```

```{r}
df <- merge(wide_fmri_data_filtered,speech_data_filtered, by = "id")

write.csv(df, file = "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/speech_and_ROI.csv", row.names = FALSE)
```

```{r}
rm(speech_data_filtered,wide_fmri_data_filtered)
```

```{r}
X <- df[, c(
  "pos_frequency_related", "pos_energy_related", "pos_spectral_balance", 
  "pos_temporal", "pos_lexical_richness", "pos_sentiment", 
  "pos_word_types", "pos_syntactic_complexity", 
  "neg_frequency_related", "neg_energy_related", "neg_spectral_balance", 
  "neg_temporal", "neg_lexical_richness", "neg_sentiment", 
  "neg_word_types", "neg_syntactic_complexity"
)]

Y2 <- df[, c("AP_aInsula_BA13", "AP_Amygdala","AP_aSTG_BA22","AP_aSTG_BA22_2",
"AP_Caudate_nucleus","AP_Cerebellum","AP_IFG_pOp_BA44","AP_IFG_pOp_BA44_2",
"AP_IFG_pOrb_BA47","AP_IFG_pOrb_BA47_2","AP_IFG_pOrb_BA47_3","AP_IFG_pTri_BA45",
"AP_MFG_BA9","AP_MFG_BA10","AP_MTG_BA21","AP_Parahippocampal_gyrus_BA28",
"AP_pSTG_BA22","AP_pSTG_BA22_2","AP_pSTG_BA22_3","AP_Putamen","AP_SMA_BA6",
"AP_SMA_BA6_3","AP_SMG_BA40","AP_SMG_BA40_7","AP_SMG_BA40_7_2","AP_Subcallosal_gyrus_BA34",
"AP_Thalamus","LP_aInsula_BA13","LP_aInsula_BA13_2","LP_Caudate_nucleus","LP_Cerebellum",
"LP_Cerebellum_2","LP_Cerebellum_3","LP_Claustrum","LP_Cuneus_BA17",
"LP_Heschls_gyrus_BA41","LP_IFG_pOp_BA44","LP_IFG_pOp_BA44_2","LP_IFG_pOrb_BA47",
"LP_Insula_BA13","LP_MFG_BA9","LP_MFG_BA9_2","LP_pSTG_BA22","LP_pSTG_BA22_2","LP_SMA_BA6", "LP_SMA_BA6_2","LP_SMG_BA40","LP_SMG_BA40_2","LP_SMG_BA40_7")]
```

```{r}
X_scaled <- scale(X) #speech
Y_scaled <- scale(Y2) #brain
```

```{r}
# Perform CCA using the cc() function from the CCA package
cca_result <- cc(X_scaled, Y_scaled)

# View the canonical correlations
print(cca_result$cor)
```

### Interpretation of the Canonical Correlations:

1.  **Understanding the Values**:

    -   The canonical correlations represent the strength of the relationship between pairs of canonical variates (linear combinations of variables) from your `X` and `Y` datasets.

    -   **Higher values** (e.g., 0.89, 0.86, 0.83) indicate a strong relationship between the canonical variates.

    -   **Lower values** (e.g., 0.29, 0.35) suggest a weaker relationship.

2.  **General Interpretation**:

    -   **Strong Canonical Correlations (\> 0.7)**: The first few canonical correlations are relatively high, which means that there are strong linear relationships between the first few pairs of canonical variates.
        These are likely the most interpretable and meaningful pairs.

    -   **Moderate Canonical Correlations (0.5 - 0.7)**: Mid-range correlations suggest moderate relationships, which may still be meaningful but warrant closer examination.

    -   **Weak Canonical Correlations (\< 0.5)**: The lower correlations indicate weaker relationships, and these pairs of canonical variates might not contribute much to the overall relationship between the datasets.

3.  **Key Points for Interpretation**:

    -   **First Pair of Canonical Variates**: The first canonical correlation (0.893) is typically the most important, representing the strongest linear relationship between the two datasets.

    -   **Diminishing Importance**: As you move to the right in the list of canonical correlations, the importance of each subsequent pair diminishes.
        Often, only the first few pairs are of practical significance.

    -   **Number of Significant Canonical Correlations**: You might want to focus on the first few canonical correlations that are significantly different from zero.
        This can be assessed using a statistical test, such as Wilks' Lambda.

### Canonical Variates:

You can examine the canonical variates and their loadings to understand which variables in your `X` and `Y` datasets are contributing the most to the relationships:

```{r}
# Canonical variates for X
X_canonical <- cca_result$xcoef

# Canonical variates for Y
Y_canonical <- cca_result$ycoef

# View the first few rows of canonical variates
head(X_canonical)
head(Y_canonical)
```

### **Number of Canonical Correlations**:

-   The number of canonical correlations is equal to the minimum number of variables in the `X` and `Y` datasets. Since you have 16 canonical correlations, it suggests that either `X` or `Y` (or both) have at least 16 variables. If one set had fewer than 16 variables, the number of canonical correlations would be limited to that smaller number.

1.  **Canonical Variates**:

    -   Each canonical correlation corresponds to a pair of canonical variates—linear combinations of the original variables in `X` and `Y`.

    -   The first pair of canonical variates corresponds to the largest canonical correlation, representing the strongest linear relationship between the two datasets.

    -   The second pair corresponds to the second-largest correlation, and so on, with each subsequent pair representing progressively weaker relationships.

2.  **Interpreting Multiple Canonical Correlations**:

    -   **First Canonical Correlation**: This is usually the most important and strongest relationship between the datasets.
        It often captures the primary pattern of association.

    -   **Subsequent Canonical Correlations**: These capture additional patterns of association, but they are usually weaker and may explain more subtle or less pronounced relationships.

    -   **Significance Testing**: Typically, not all canonical correlations are statistically significant.
        Higher-order canonical correlations (those with smaller values) might not be significantly different from zero, meaning they don't represent meaningful relationships between the datasets.

### What to Do with the 16 Canonical Correlations:

1.  **Focus on Significant Correlations**:

    -   After performing CCA, you would normally assess which of these correlations are statistically significant.
        Often, only the first few canonical correlations are significant and meaningful for interpretation.

    -   Use a Wilks' Lambda test or similar to determine which canonical correlations are significant.

2.  **Interpret the First Few Canonical Correlations**:

    -   The first few canonical correlations (typically 1st and 2nd) usually capture the most interpretable and strongest relationships between the speech features and brain activities.

    -   For these, examine the canonical variates to see which specific variables in `X` and `Y` are driving these relationships.

3.  **Ignore Non-Significant Correlations**:

    -   If some of the canonical correlations are not significant, they can generally be ignored in the interpretation. These correlations might represent noise rather than meaningful relationships.

## Test which correlations are significant

Wilks' Lambda test

```{r}
# Calculate the p-values for the canonical correlations
p_values <- p.asym(cca_result$cor, nrow(X_scaled), ncol(X_scaled), ncol(Y_scaled))
print(p_values)
```

```{r}
# Create a data frame to store the results
canonical_significance <- data.frame(
  Canonical_Variates = paste0(1:length(p_values$stat), " to 16"),
  Wilks_Stat = p_values$stat,
  Approx_F = p_values$approx,
  DF1 = p_values$df1,
  DF2 = p_values$df2,
  P_Value = p_values$p.value
)
write.csv(canonical_significance, file = "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/canonical_significance_results.csv", row.names = FALSE)
```

**12th to 16th cannonical**: The p-values are above 0.05, with values like p=0.12p = 0.12p=0.12, p=0.495p = 0.495p=0.495, etc., indicating these correlations are not statistically significant.

The first 9 canonical correlations are statistically significant, and these are the pairs of canonical variates that explain meaningful relationships between your speech features and brain region activities.

Focus on interpreting the canonical variates for the first few significant correlations.
Look at the canonical loadings (coefficients) to determine which variables in `X` and `Y` are contributing most to these relationships.

## Get variance explained and correlation for each Canonical

```{r}
canonical_correlations <- cca_result$cor
variance_explained <- canonical_correlations^2
canonical_table <- data.frame(
  Canonical_Variate = paste0("Canonical ", 1:length(canonical_correlations)),
  Canonical_Correlation = round(canonical_correlations, 3),
  Variance_Explained = round(variance_explained, 3)
)
write.csv(canonical_table, "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/canonical_correlations_and_variance.csv", row.names = FALSE)

```

# Extract and Interpret the First four Canonical Variate:

Extract canonical scores

```{r}
# 'cca_result' is the result from your CCA

# Compute canonical scores for X and Y variables
X_scores <- X_scaled %*% cca_result$xcoef
Y_scores <- Y_scaled %*% cca_result$ycoef

# Convert to data frames for easier manipulation
X_scores_df <- as.data.frame(X_scores)
Y_scores_df <- as.data.frame(Y_scores)

# Name the canonical scores for clarity (e.g., Canonical1, Canonical2, ...)
colnames(X_scores_df) <- paste0("Canonical_Speech", 1:ncol(X_scores_df))
colnames(Y_scores_df) <- paste0("Canonical_Brain", 1:ncol(Y_scores_df))

# Subset only the first six canonical variates
X_scores_df <- X_scores_df[, 1:4]
Y_scores_df <- Y_scores_df[, 1:4]
```

Interpret Canonical Loadings

```{r}
# Assuming cca_result is your object with canonical correlation analysis results
# Extract the canonical loadings for X (speech features) and Y (brain regions)
canonical_loadings_X <- cca_result$xcoef
canonical_loadings_Y <- cca_result$ycoef

# Convert to data frames for easier manipulation
canonical_loadings_X_df <- as.data.frame(canonical_loadings_X)
canonical_loadings_Y_df <- as.data.frame(canonical_loadings_Y)

# Name the columns to represent the canonical variates
colnames(canonical_loadings_X_df) <- paste0("Canonical_", 1:ncol(canonical_loadings_X_df))
colnames(canonical_loadings_Y_df) <- paste0("Canonical_", 1:ncol(canonical_loadings_Y_df))

# Add the variable names (assuming you have rownames for your speech and brain variables)
canonical_loadings_X_df$Variable <- rownames(canonical_loadings_X_df)
canonical_loadings_Y_df$Variable <- rownames(canonical_loadings_Y_df)

# Add dataset labels (Speech or Brain)
canonical_loadings_X_df$Dataset <- "Speech"
canonical_loadings_Y_df$Dataset <- "Brain"

```

```{r}
# Combine the canonical loadings for X and Y into a single data frame
canonical_loadings_combined <- data.frame(
  Variable = c(rownames(canonical_loadings_X_df), rownames(canonical_loadings_Y_df)),
  Dataset = c(rep("Speech", nrow(canonical_loadings_X_df)), rep("Brain", nrow(canonical_loadings_Y_df))),
  Canonical_1 = c(canonical_loadings_X_df[,1], canonical_loadings_Y_df[,1]),
  Canonical_2 = c(canonical_loadings_X_df[,2], canonical_loadings_Y_df[,2]),
  Canonical_3 = c(canonical_loadings_X_df[,3], canonical_loadings_Y_df[,3]),
  Canonical_4 = c(canonical_loadings_X_df[,4], canonical_loadings_Y_df[,4])
)


head(canonical_loadings_combined)
write.csv(canonical_loadings_combined,
          "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/canonical_loadings_combined.csv", row.names = FALSE)

```

Get the top loading features for the canonical

```{r}
# Function to extract top 3 contributors for each canonical variate from the speech and brain datasets
get_top_contributors <- function(canonical_data, canonical_var, dataset_label, top_n = 3) {
  # Sort by absolute values and take top_n
  top_vars <- canonical_data %>%
    arrange(desc(abs(.data[[canonical_var]]))) %>%
    slice(1:top_n) %>%
    mutate(Dataset = dataset_label, Canonical_Variate = canonical_var) %>%
    select(Variable, Dataset, Canonical_Variate, Loading = .data[[canonical_var]])
  
  return(top_vars)
}

# Prepare an empty data frame to store results
top_contributors <- data.frame()

# Loop over the first 6 canonical variates (using your column names Canonical_1, Canonical_2, etc.)
for (i in 1:6) {
  canonical_var <- paste0("Canonical_", i)  # Your column names are Canonical_1, Canonical_2, etc.
  
  # Get top 3 brain regions
  top_brain <- get_top_contributors(canonical_loadings_Y_df, canonical_var, "Brain")
  
  # Get top 3 speech features
  top_speech <- get_top_contributors(canonical_loadings_X_df, canonical_var, "Speech")
  
  # Combine brain and speech top variables for this canonical variate
  top_combined <- rbind(top_brain, top_speech)
  
  # Add to the main data frame
  top_contributors <- rbind(top_contributors, top_combined)
}

# Pivot the data into the desired wide format
top_contributors_wide <- top_contributors %>%
  pivot_wider(names_from = Canonical_Variate, values_from = Loading, names_prefix = "Canonical_")

# Save the final table to CSV
write.csv(top_contributors_wide, "W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/mri_speech/top_contributors_first_6_canonicals.csv", row.names = FALSE)

# View the table in the console
print(top_contributors_wide)
```
