---
title: "Subjective Rating"
output: html_notebook
---

## Import Libraries

```{r}
library(readxl)
library(readr)
library(dplyr)
library(ggplot2)
library(writexl)
library(tidyr)
library(purrr)  # Load the purrr package for the reduce function
library(stringr)
library(grid)
library(cowplot)
library(ggprism)
library(lme4)
library(lmerTest)
library(lm.beta)
#The car package can provide Type III ANOVA tables with p-values for mixed models. 
library(car)
library(emmeans)
# Create APA Style Tables
library(knitr)
# To export to office
library(officer)
library(prism)
library(afex) 
```

## Slider Rating Analysis

```{r}
df <- read_csv("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/slider_rating.csv",col_types = cols(Rating = col_number(),Count = col_number()))
```

Include information about age and sex

```{r}
Demographics <- read_excel("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/Demographics.xlsx",col_types = c("text", "numeric", "numeric"))
```

```{r}
# Rename the 'bids_number' column in df_age_sex to match 'participantID' in df_ratings
colnames(Demographics)[which(names(Demographics) == "Bids-Nummer")] <- "participantID"
# Merge the data frames by 'participantID'
df <- merge(df, Demographics, by = "participantID")
head(df)
```

```{r}
df$group <- as.factor(df$group)
df$condition <- as.factor(df$condition)
df$participantID <- as.factor(df$participantID)
df$sex <- as.factor(df$Gender_f1_m2)
df$age <- as.numeric(df$Alter)
```

```{r}
# Recalculate mean and standard deviation for each participant
participant_aggregates <- df %>%
  group_by(participantID, group, condition, age, sex) %>%
  summarize(
    mean_rating = sum(Rating * Count) / sum(Count),
    sd_rating = sqrt(sum(Count * (Rating - mean_rating)^2) / sum(Count)),
    n_trials = sum(Count),
    .groups = 'drop'
  )
print(participant_aggregates)
```

### Descriptive Statistics

```{r}
summary_stats <- participant_aggregates %>%
  group_by(group, condition) %>%
  summarize(
    mean_rating = mean(mean_rating, na.rm = TRUE),
    sd_rating = mean(sd_rating, na.rm = TRUE),
    n = n()
  )
print(summary_stats)
```

### Group differences

#### Visual Data Inspection

```{r}
# Histogram of the dependent variable
ggplot(participant_aggregates, aes(x = mean_rating)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  labs(title = "Histogram of Mean Rating", x = "Mean Rating", y = "Count")

# Scatterplot of independent variables vs. dependent variable
ggplot(participant_aggregates, aes(x = group, y = mean_rating)) +
  geom_jitter(width = 0.2, height = 0.2) +
  labs(title = "Scatterplot of Group vs. Mean Rating", x = "Group", y = "Mean Rating")
```

**Two Peaks**: A bimodal distribution has two distinct peaks or clusters.
For example, in your case, one peak might be around a rating of 3, and another around 6.

But if we dont separate our data into two categories, we can still use gaussian distribution

## Estimate Model

```{r}
model <- lmer(mean_rating ~ group*condition + age + sex + (1|participantID),
                   data = participant_aggregates)

# Summary of the model
model_anova <- anova(model, type = 3)
model_anova
```

#### Post Hoc Tests

```{r}
eff.group       <- emmeans(model,pairwise ~ group,adjust ="bonferroni")
eff.condition   <- emmeans(model,pairwise ~ condition,adjust ="bonferroni")
eff.interaction <- emmeans(model,pairwise ~ condition:group,adjust ="bonferroni")
```

```{r}
# Extract the estimated marginal means
emm_group       <- as.data.frame(eff.group$emmeans)
emm_condition   <- as.data.frame(eff.condition$emmeans)
emm_interaction <- as.data.frame(eff.interaction$emmeans)

# Extract the pairwise comparisons with p-values
pairs_group       <- as.data.frame(eff.group$contrasts)
pairs_condition   <- as.data.frame(eff.condition$contrasts)
pairs_interaction <- as.data.frame(eff.interaction$contrasts)
```

### Export Results

#### Anova

```{r}
# Assuming your ANOVA results are stored in 'results' and it's a data frame
anova_df <- as.data.frame(model_anova)

# Add asterisks for significance levels
anova_df <- anova_df %>%
  mutate(asterisks = case_when(
    `Pr(>F)` < 0.001 ~ "***",
    `Pr(>F)` < 0.01  ~ "**",
    `Pr(>F)` < 0.05  ~ "*",
    TRUE ~ ""
  ))

anova_df
# Select and rename the columns for APA style
apa_table <- anova_df %>%
  select(`Sum Sq`, `Mean Sq`, `F value`, `Pr(>F)`, `asterisks`, NumDF, DenDF) %>%
  rename(
    `Num df` = NumDF, # Numerator dr, number of independent comparisons or groups being tested
    `Den df` = DenDF, # amount of independent information in the data, df of the error term
    `Sum of Squares` = `Sum Sq`,
    `Mean Square` = `Mean Sq`,
    `F` = `F value`,
    `p` = `Pr(>F)`,
    `Significance` = asterisks
    ) %>% 
  mutate(
    `Sum of Squares` = round(`Sum of Squares`, 2),
    `Mean Square` = round(`Mean Square`, 2),
    `Num df` = round(`Num df`, 2),      # Round Num df to 2 decimal places
    `Den df` = round(`Den df`, 2),       # Round Den df to 2 decimal places
    `F` = round(`F`, 2),
    `p` = round(`p`, 3)    # Round p-values to 2 decimal places

  )

# Create a new Word document
doc <- read_docx()
# Add a title to the document
doc <- doc %>%
  body_add_table(value = apa_table, style = "table_template")  
# Save the document
print(doc, target = "glmm_subjective_rating_Anova.docx")
```

#### Group

```{r}
pairs_group_df <- as.data.frame(pairs_group)
apa_table <- pairs_group_df %>%
  select(contrast, estimate, SE, df, t.ratio, p.value) %>%
  rename(
    `Comparison` = contrast,
    `Mean Difference` = estimate,
    `Standard Error` = SE,
    `Degrees of Freedom` = df,
    `t-Value` = t.ratio,
    `p-Value` = p.value)


apa_table$p.Value <- format(apa_table$p.Value, scientific = FALSE)

doc <- read_docx()
doc <- doc %>%
  body_add_par("Post-Hoc Comparisons Between Conditions", style = "heading 1") %>% 
  body_add_table(value = apa_table, style = "table_template")

print(doc, target = "subjective_rating_Group_Post_Hoc.docx")
```

#### Condition

```{r}
str(pairs_condition_df)
```

```{r}
pairs_condition_df <- as.data.frame(pairs_condition)
apa_table <- pairs_condition_df %>%
  select(contrast, estimate, SE, df, t.ratio, p.value) %>%
  rename(
    `Comparison` = contrast,
    `Mean Difference` = estimate,
    `t` = t.ratio,
    `p` = p.value
  ) %>% 
  mutate(
    `SE` = round(SE, 2),
    `df` = round(df, 2),
    `Mean Difference` = round(`Mean Difference`, 2),
    `t` = round(`t`, 2),
    `p` = round(`p`, 3)
  )


doc <- read_docx()
doc <- doc %>%
  body_add_par("Post-Hoc Comparisons Between Conditions", style = "heading 1") %>% 
  body_add_table(value = apa_table, style = "table_template")
print(doc, target = "subjective_rating_Condition_Post_Hoc.docx")

```

#### Interaction Group\*Condition

```{r}
pairs_interaction_df <- as.data.frame(pairs_interaction)
apa_table <- pairs_interaction_df %>%
  select(contrast, estimate, SE, df, t.ratio, p.value) %>%
  rename(
    `Comparison` = contrast,
    `Mean Difference` = estimate,
    `t` = t.ratio,
    `p` = p.value
  ) %>% 
  mutate(
    `SE` = round(`SE`, 2),
    `df` = round(`df`, 2),
    `Mean Difference` = round(`Mean Difference`, 2),
    `t` = round(`t`, 2),
    `p` = round(`p`, 3)
  )

doc <- read_docx()
doc <- doc %>%
  body_add_par("Post-Hoc Comparisons Between Conditions", style = "heading 1") %>% 
  body_add_table(value = apa_table, style = "table_template")

print(doc, target = "subjective_rating_Interaction_Post_Hoc.docx")
```

### Plot Results

```{r}
participant_aggregates$condition <- factor(participant_aggregates$condition, 
                                           levels = c("happy_happy", 
                                                      "sad_happy", 
                                                      "happy_sad", 
                                                      "sad_sad"))
```

```{r}
group <- afex_plot(model, x = "group", 
            id = "participantID", dodge = 0.8, 
            data_geom = geom_violin, 
            mapping = c("linetype", "fill"),
            error_arg = list(width = 0.2),
            legend_title = "cluster number",
            data_arg = list(width = 0.5)) + 
    scale_fill_manual(values = c("HC" = "turquoise", 
                                 "MDD" = "violet")) + 
    theme_prism() +
    theme(legend.position = "none")+
    labs(y = "Rating from sad (0) to happy (9)", x = "")+
  ylim(0,9)
  
conditions <-  afex_plot(model, x = "condition", 
                         id = "participantID", dodge = 0.8,
                         data_geom = geom_violin, 
                         mapping = "fill",
                         error_arg = list(width = 0.2),
                         data_arg = list(width = 0.5)) +
  scale_fill_manual(values = c("happy_happy" = "darkgreen", 
                               "sad_happy" = "lightgreen",
                               "happy_sad" ="lightblue", 
                               "sad_sad" ="darkblue")) +
  theme_prism() +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 10))+
  labs(title = "", y = "Rating from sad (0) to happy (9)", x = "")+
  ylim(0,9)

con_in_group <- afex_plot(model, x = "condition", panel = "group", 
                          id = "participantID", dodge = 0.8,
                          data_geom = geom_violin, 
                          mapping = c("linetype", "fill", "group"), 
                          error_arg = list(width = 0.2),
                          legend_title = "",
                          data_arg = list(width = 0.5)) + 
  theme_prism() +
  theme(legend.position = "none")+
  scale_fill_manual(values = c("happy_happy" = "darkgreen",
                               "sad_happy" = "lightgreen",
                               "happy_sad" ="lightblue", 
                               "sad_sad" ="darkblue")) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 10))+
  labs(y = "Rating from sad (0) to happy (9)", x = "")+
  ylim(0,9)
```

```{r, fig.width=12, fig.height=4.5}
combined_plots <- cowplot::plot_grid(group,conditions,con_in_group,
                                     labels = c("A", "B", "C"),
                                     ncol = 3,nrow = 1,
                                     rel_heights = c(1,1,1),
                                     rel_widths = c(0.7,1,2),
                                     axis = "tb",# top and bottom axes should be aligned.
                                     align = "h")
print(combined_plots)
ggsave("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/plots/glmm_subjective_rating1.tiff", combined_plots, width = 12, height = 4.5)
```
