---
title: "GLMM"
output: html_notebook
---

Import modules

```{r}
library(lme4)         # mixed model package
library(lmerTest)     # library providing p-values for mixed models in lme4
library(readxl)       # read excel
library(ggplot2)      # graphics
library(emmeans)      # library for post-hoc tests
library(pbkrtest)     # needed for post-hoc tests in mixed models
library(jtools)       # post hoc tests
library(interactions) 
library(effects)      # estimate effects
library(apaTables)
library(dplyr)
library(tidyr)
library(effects)
library(readxl)
library(readr)
library(fitdistrplus)
library(ggplot2)
library(openxlsx)
library(afex) 
library(cowplot)
library(prism)
```

# RT Analysis

## Read in dataset

```{r}
df <- read.csv("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/ratings_by_condition_and_participant.csv")
# Assuming your dataframe is named 'df'
clean_df <- na.omit(df)
```

```{r}
df <- clean_df%>%
  mutate(sex = as.factor(sex)) %>%
  mutate(group = as.factor(group)) %>%
  mutate(condition = as.factor(condition)) %>%
  mutate(age = as.numeric(age))%>%
  mutate(rt = as.numeric(key_resp.rt))%>%
  mutate(accuracy = as.numeric(correct))
# Ensure that your 'group' variable is a factor with correct labels
df$group <- factor(df$group, levels = c(1, 2), labels = c("MDD", "HC"))

```

## Select distribution

```{r}
fit_gamma <- fitdist(df$key_resp.rt, "gamma")
fit_norm <- fitdist(df$key_resp.rt, "norm")

```

```{r}
plot_fits <- function(df, fit_gamma, fit_norm) {
  ggplot(df, aes(x = key_resp.rt)) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black", alpha = 0.6) +
    stat_function(aes(color = "Gamma"), fun = dgamma, args = list(shape = fit_gamma$estimate["shape"], rate = fit_gamma$estimate["rate"]),
                  size = 1, linetype = "dashed") +
    stat_function(aes(color = "Normal"), fun = dnorm, args = list(mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"]),
                  size = 1, linetype = "solid") +
    labs(title = "Response Time Distribution Fit",
         x = "Response Time",
         y = "Density",
         color = "Distribution") +
    theme_minimal() +
    scale_color_manual(values = c("Gamma" = "green", "Normal" = "red"))
}

plot_fits(df, fit_gamma, fit_norm)

```

```{r}
gof_stat <- gofstat(list(fit_gamma, fit_norm))
print(gof_stat)

```

Based on the provided goodness-of-fit statistics and criteria, we can interpret the fit of the gamma and normal distributions to your data as follows:

### Goodness-of-fit Statistics

1.  **Kolmogorov-Smirnov (KS) Statistic:**
    -   Gamma: 0.03380856
    -   Normal: 0.0438622
    -   Interpretation: The gamma distribution has a lower KS statistic, indicating a better fit to the data compared to the normal distribution.
2.  **Cramer-von Mises Statistic:**
    -   Gamma: 0.04337989
    -   Normal: 0.2003313
    -   Interpretation: The gamma distribution has a significantly lower Cramer-von Mises statistic, suggesting a much better fit than the normal distribution.
3.  **Anderson-Darling Statistic:**
    -   Gamma: 0.32097548
    -   Normal: 1.0736581
    -   Interpretation: The gamma distribution has a lower Anderson-Darling statistic, indicating a better fit compared to the normal distribution.

### Goodness-of-fit Criteria

1.  **Akaike's Information Criterion (AIC):**
    -   Gamma: -333.4991
    -   Normal: -323.0050
    -   Interpretation: The gamma distribution has a lower AIC, which means it is a better fit to the data while penalizing for the number of parameters.
2.  **Bayesian Information Criterion (BIC):**
    -   Gamma: -325.1936
    -   Normal: -314.6995
    -   Interpretation: The gamma distribution also has a lower BIC, reinforcing the conclusion that it fits the data better than the normal distribution.

### Overall Conclusion

Based on all the provided goodness-of-fit statistics and criteria (KS statistic, Cramer-von Mises statistic, Anderson-Darling statistic, AIC, and BIC), the gamma distribution fits your data better than the normal distribution.
This conclusion is drawn from the fact that all the metrics show lower values for the gamma distribution, indicating a closer match to your observed data.

## Estimate GLMM

```{r}

model1 <-lmer(df$key_resp.rt~ condition*group + age + sex + (1|participantID),data = df)#, family = gaussian())
anova_Model<-anova(model1, type = 3, ddf= "Kenward-Roger")
```

```{r}
anova_Model
```

```{r}
#sjPlot::tab_model(model1)
#sjPlot::plot_model(model1, type = "diag")
```

```{r}
# Convert ANOVA results to a data frame
anova_df <- broom::tidy(anova_Model)

# View the table
print(anova_df)

# Save the table to a CSV file
write.csv(anova_df, "glmm_rt_anova_results.csv", row.names = FALSE)
```

### Post Hoc Tests

```{r}
eff.group       <- emmeans(model1, pairwise ~ group)
eff.condition   <- emmeans(model1, pairwise ~ condition,adjust ="bonferroni")
```

```{r}
# Plot histograms to check the distribution
hist(df$age, main = "Age Distribution", xlab = "Age")
hist(df$key_resp.rt, main = "Response Time Distribution", xlab = "Response Time")

# Plot Q-Q plots to check normality
qqnorm(df$age); qqline(df$age, col = "red")
qqnorm(df$key_resp.rt); qqline(df$key_resp.rt, col = "red")

```

Use Spearman correlation for non-normal distributed data

```{r}
# Spearman correlation
eff.age <- cor.test(df$age, df$key_resp.rt, method = "spearman")

```

```{r}
print(eff.age )
print(eff.condition)
print(eff.group)
```

```{r}
# Extract the summary of the emmeans results for group
group_summary <- summary(eff.group$emmeans)
group_contrasts <- summary(eff.group$contrasts)

# Convert to data frames
group_summary_df <- as.data.frame(group_summary)
group_contrasts_df <- as.data.frame(group_contrasts)

# Add a column to identify the type of results
group_summary_df$type <- "group_emmeans"
group_contrasts_df$type <- "group_contrasts"

# Extract the summary of the emmeans results for condition
condition_summary <- summary(eff.condition$emmeans)
condition_contrasts <- summary(eff.condition$contrasts)

# Convert to data frames
condition_summary_df <- as.data.frame(condition_summary)
condition_contrasts_df <- as.data.frame(condition_contrasts)

# Add a column to identify the type of results
condition_summary_df$type <- "condition_emmeans"
condition_contrasts_df$type <- "condition_contrasts"

```

```{r}
# Use Spearman correlation for non-normal distributed data
eff.age <- cor.test(df$age, df$key_resp.rt, method = "spearman")

# Extract correlation results into a data frame
age_correlation_df <- data.frame(
  term = "age_correlation",
  statistic = eff.age$statistic,
  p.value = eff.age$p.value,
  estimate = eff.age$estimate,
  method = eff.age$method,
  alternative = eff.age$alternative,
  type = "age_correlation"
)

```

```{r}
# Combine all data frames into one
combined_results <- bind_rows(
  group_summary_df,
  group_contrasts_df,
  condition_summary_df,
  condition_contrasts_df,
  age_correlation_df
)

```

```{r}
# Write to an Excel file
write.xlsx(combined_results, "glmm_rt_post_hoc_results.xlsx", rowNames = FALSE)
```

# Plot results

```{r}

# Reorder the levels of the 'condition' variable
df$condition <- factor(df$condition, levels = c("happy_happy", "sad_happy", "happy_sad", "sad_sad"))
group <-  afex_plot(model1, x = "group", 
            id = "participantID", dodge = 0.8, 
            data_geom = geom_violin, 
            mapping = c("linetype", "fill"),
            error_arg = list(width = 0.2),
            legend_title = "cluster number",
            data_arg = list(width = 0.5)) + 
    #ggpubr::theme_pubr() + 
    scale_fill_manual(values = c("HC" = "turquoise", 
                                 "MDD" = "violet")) + 
    theme_prism() +
    theme(legend.position = "none")+
    labs(y = "Reaction Time", x = "")+
  ylim(0,1.5)
  
conditions <-  afex_plot(model1, x = "condition", 
            id = "participantID", dodge = 0.8,
            data_geom = geom_violin, 
            mapping = "fill",
            error_arg = list(width = 0.2),
            data_arg = list(width = 0.5)) + #ggpubr::theme_pubr() + 
    scale_fill_manual(values = c("happy_happy" = "darkgreen", 
                                 "sad_happy" = "lightgreen",
                                 "happy_sad" ="lightblue", 
                                 "sad_sad" ="darkblue")) +
    theme_prism() +
  theme(legend.position = "none")+
    theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 10))+
    labs(y = "Reaction Time", x = "")+
  ylim(0,1.5)

#colors()

age <- ggplot(df, aes(x = age, y = key_resp.rt)) +
  geom_point(alpha = 0.6, color = "#1f77b4") + 
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +  
  labs(x = "Age",
       y = "Reaction Time") +
  theme_prism()+
  ylim(0,1.5)

```

```{r, fig.width=10, fig.height=5}
combined_plots <- cowplot::plot_grid(group,age,conditions,
                            labels = c("A", "B", "C"),
                            ncol = 3,nrow = 1,
                            rel_heights = c(1,1),
                            align = "h")
print(combined_plots)
```

```{r}
ggsave("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/plots/glmm_rt.tiff", combined_plots, width = 10, height = 5)

```

```{r}
library(dplyr)

# Count unique participants in each group
df %>% 
  group_by(group) %>% 
  summarise(Num_Participants = n_distinct(participantID))
```
