---
title: "GLMM"
output: html_notebook
---

Import modules

```{r}
library(lme4)         # mixed model package
library(lmerTest)     # library providing p-values for mixed models in lme4
library(readxl)       # read excel
library(ggplot2)      # graphics
library(emmeans)      # library for post-hoc tests
library(pbkrtest)     # needed for post-hoc tests in mixed models
library(jtools)       # post hoc tests
library(interactions) 
library(effects)      # estimate effects
library(apaTables)
library(dplyr)
library(tidyr)
library(readr)
library(fitdistrplus)
library(openxlsx)
library(afex) 
library(cowplot)
library(prism)
```

# Accuracy Analysis

## Read in dataset

```{r}
df <- read.csv("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/data/ratings_by_condition_and_participant.csv")
# Assuming your dataframe is named 'df'
clean_df <- na.omit(df)
```

```{r}
df <- clean_df%>%
  mutate(sex = as.factor(sex)) %>%
  mutate(group = as.factor(group)) %>%
  mutate(condition = as.factor(condition)) %>%
  mutate(age = as.numeric(age))%>%
  mutate(rt = as.numeric(key_resp.rt))%>%
  mutate(accuracy = as.numeric(correct))
# Ensure that your 'group' variable is a factor with correct labels
df$group <- factor(df$group, levels = c(1, 2), labels = c("MDD", "HC"))

```

## Select distribution

```{r}
data <- df
x<-data$accuracy +100

# Fit different distributions
fit_gamma <- fitdist(x, "gamma")
fit_norm <- fitdist(x, "norm")

plot_fits <- function(df, fit_gamma, fit_norm) {
  ggplot(df, aes(x = x)) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black", alpha = 0.6) +
    stat_function(aes(color = "Gamma"), fun = dgamma, args = list(shape = fit_gamma$estimate["shape"], rate = fit_gamma$estimate["rate"]),
                  size = 1, linetype = "dashed") +
    stat_function(aes(color = "Normal"), fun = dnorm, args = list(mean = fit_norm$estimate["mean"], sd = fit_norm$estimate["sd"]),
                  size = 1, linetype = "solid") +
    labs(title = "Accuracy Distribution Fit",
         x = "accuracy",
         y = "Density",
         color = "Distribution") +
    theme_minimal() +
    scale_color_manual(values = c("Gamma" = "green", "Normal" = "red"))
}

plot_fits(df, fit_gamma, fit_norm)
```

```{r}
gof_stat <- gofstat(list(fit_gamma, fit_norm))
print(gof_stat)
```

### Goodness-of-Fit Statistics

1.  **Kolmogorov-Smirnov Statistic**: Measures the maximum difference between the empirical distribution function and the cumulative distribution function of the fitted distribution.
    Lower values indicate a better fit.

    -   Gamma: 0.2650154

    -   Normal: 0.249860

    -   **Interpretation**: The normal distribution fits the data slightly better than the gamma distribution according to this statistic.

2.  **Cramer-von Mises Statistic**: Measures the squared differences between the empirical and fitted cumulative distribution functions.
    Lower values indicate a better fit.

    -   Gamma: 9.1876182

    -   Normal: 7.862238

    -   **Interpretation**: The normal distribution fits the data better than the gamma distribution according to this statistic.

3.  **Anderson-Darling Statistic**: Similar to the Cramer-von Mises statistic but gives more weight to the tails of the distribution.
    Lower values indicate a better fit.

    -   Gamma: 50.5817331

    -   Normal: 44.358003

    -   **Interpretation**: The normal distribution fits the data better than the gamma distribution according to this statistic.

### Goodness-of-Fit Criteria

1.  **Akaike's Information Criterion (AIC)**: Measures the relative quality of statistical models for a given set of data.
    Lower values indicate a better fit.

    -   Gamma: 4198.028

    -   Normal: 4097.460

    -   **Interpretation**: The normal distribution has a lower AIC, indicating a better fit compared to the gamma distribution.

2.  **Bayesian Information Criterion (BIC)**: Similar to AIC but includes a penalty term for the number of parameters in the model.
    Lower values indicate a better fit.

    -   Gamma: 4206.334

    -   Normal: 4105.766

    -   **Interpretation**: The normal distribution has a lower BIC, indicating a better fit compared to the gamma distribution.

### Conclusion

Based on the provided statistics and criteria, the normal distribution fits your data better than the gamma distribution.

## Estimate Model

```{r}
model <-lmer(df$accuracy~ condition*group + age + sex + (1|participantID),data = df)
anova_Model<-anova(model, type = 3, ddf= "Kenward-Roger")
```

```{r}
anova_Model
```

```{r}
#sjPlot::tab_model(model1)
#sjPlot::plot_model(model1, type = "diag")
```

```{r}
# Convert ANOVA results to a data frame
anova_df <- broom::tidy(anova_Model)

# View the table
print(anova_df)

# Save the table to a CSV file
write.csv(anova_df, "glmm_accuracy_anova_results.csv", row.names = FALSE)
```

### Post Hoc Tests

```{r}
eff.group_cond  <- emmeans(model, pairwise ~ group*condition, adjust ="bonferroni")
eff.condition   <- emmeans(model, pairwise ~ condition,adjust ="bonferroni")
```

```{r}
print(eff.condition)
print(eff.group_cond)
```

```{r}
# Extract the summary of the emmeans results for group
group_summary <- summary(eff.group_cond$emmeans)
group_contrasts <- summary(eff.group_cond$contrasts)

# Convert to data frames
group_summary_df <- as.data.frame(group_summary)
group_contrasts_df <- as.data.frame(group_contrasts)

# Add a column to identify the type of results
group_summary_df$type <- "group_emmeans"
group_contrasts_df$type <- "group_contrasts"

# Extract the summary of the emmeans results for condition
condition_summary <- summary(eff.condition$emmeans)
condition_contrasts <- summary(eff.condition$contrasts)

# Convert to data frames
condition_summary_df <- as.data.frame(condition_summary)
condition_contrasts_df <- as.data.frame(condition_contrasts)

# Add a column to identify the type of results
condition_summary_df$type <- "condition_emmeans"
condition_contrasts_df$type <- "condition_contrasts"

```

```{r}
# Combine all data frames into one
combined_results <- bind_rows(
  group_summary_df,
  group_contrasts_df,
  condition_summary_df,
  condition_contrasts_df
)
```

```{r}
# Write to an Excel file
write.xlsx(combined_results, "glmm_accuracy_post_hoc_results.xlsx", rowNames = FALSE)
```

### Plot

```{r}

# Reorder the levels of the 'condition' variable
df$condition <- factor(df$condition, levels = c("happy_happy", "sad_happy", "happy_sad", "sad_sad"))



con_in_group <- afex_plot(model, x = "condition", panel = "group", 
                          id = "participantID", dodge = 0.8,
                          data_geom = geom_violin, 
                          mapping = c("linetype", "fill", "group"), 
                          error_arg = list(width = 0.2),
                          legend_title = "",
                          data_arg = list(width = 0.5)) + 
  theme_prism() +
  theme(legend.position = "none")+
  scale_fill_manual(values = c("happy_happy" = "darkgreen",
                               "sad_happy" = "lightgreen",
                               "happy_sad" ="lightblue", 
                               "sad_sad" ="darkblue")) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 10))+
  labs(y = "Accuracy", x = "")
  
conditions <-  afex_plot(model, x = "condition", 
            id = "participantID", dodge = 0.8,
            data_geom = geom_violin, 
            mapping = "fill",
            error_arg = list(width = 0.2),
            data_arg = list(width = 0.5)) + #ggpubr::theme_pubr() + 
    scale_fill_manual(values = c("happy_happy" = "darkgreen", 
                                 "sad_happy" = "lightgreen",
                                 "happy_sad" ="lightblue", 
                                 "sad_sad" ="darkblue")) +
    theme_prism() +
  theme(legend.position = "none")+
    theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1,
                                   vjust = 1,
                                   margin = margin(t = 0, b = 0),size = 10))+
    labs(title = "", y = "Accuracy", x = "")

#colors()
```

```{r, fig.width=10, fig.height=4}
combined_plots <- cowplot::plot_grid(conditions,con_in_group,
                            labels = c("A", "B"),
                            ncol = 2)
combined_plots
```

```{r}
#ggsave("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/plots/glmm_accuracy_1.tiff", conditions, width = 4, height = 4)
```

```{r}
ggsave("W:/Fmri_Forschung/Allerlei/JuliaS/GitHub/SubliminalVideoPriming/plots/glmm_accuracy_2.tiff", combined_plots, width = 10, height = 4)
```

```{r}
library(dplyr)

# Count unique participants in each group
df %>% 
  group_by(group) %>% 
  summarise(Num_Participants = n_distinct(participantID))

```

## Test Lennard

plot effects from model

```{r}
# Assuming 'model' is your fitted GLMM model and 'df' is your original dataset

# Generate predictions from the model
df$predicted_values <- predict(model, type = "response")

# Plot the predicted values using ggplot
plot1 <- ggplot(df, 
       aes(x = condition, 
           y = predicted_values, 
           fill = condition)) +
  geom_boxplot(width = 0.5, 
               outlier.shape = NA) +  # Boxplot of predicted values
  geom_jitter(aes(y = predicted_values), # Jitter points of predicted values 
              position = position_jitter(width = 0.1), 
              size = 1.5, 
              alpha = 0.3) +  # adjust transparency
  scale_fill_manual(values = c("happy_happy" = "darkgreen",
                               "sad_happy" = "lightgreen",
                               "happy_sad" = "lightblue", 
                               "sad_sad" = "darkblue")) +
  theme_prism() +
  theme(legend.position = "none", # auskommentieren, wenn du Legende haben willst
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   vjust = 1, 
                                   size = 10)) +
  labs(title = "Predicted Values", y = "Predicted Accuracy", x = "") # titel anpassen



plot2 <- ggplot(df, 
       aes(x = condition, 
           y = predicted_values, 
           fill = condition)) +
  geom_boxplot(width = 0.5, 
               outlier.shape = NA) +  
  geom_jitter(aes(y = predicted_values), 
              position = position_jitter(width = 0.1), 
              size = 1.5, 
              alpha = 0.3) +  
  scale_fill_manual(values = c("happy_happy" = "darkgreen",
                               "sad_happy" = "lightgreen",
                               "happy_sad" = "lightblue", 
                               "sad_sad" = "darkblue")) +
  theme_prism() +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   vjust = 1, 
                                   size = 10)) +
  labs(title = "Predicted Values", y = "Predicted Accuracy", x = "") 




plot3 <- ggplot(df, 
       aes(x = condition, 
           y = predicted_values, 
           fill = condition)) +
  geom_boxplot(width = 0.5, 
               outlier.shape = NA) +  
  geom_jitter(aes(y = predicted_values), 
              position = position_jitter(width = 0.1), 
              size = 1.5, 
              alpha = 0.3) + 
  scale_fill_manual(values = c("happy_happy" = "darkgreen",
                               "sad_happy" = "lightgreen",
                               "happy_sad" = "lightblue", 
                               "sad_sad" = "darkblue")) +
  theme_prism() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   vjust = 1, 
                                   size = 10)) +
  labs(title = "Predicted Values", y = "Predicted Accuracy", x = "") 


```

Kombiniere Plots

ändere figure size am anfang fig.width=10, fig.height=4

```{r, fig.width=10, fig.height=4}
#install.packages("gridExtra")
#library(gridExtra)

# Assuming you have three plots: plot1, plot2, plot3
combined_plot <- grid.arrange(plot1, plot2, plot3, ncol = 3)

# Display the combined plot
combined_plot

```
