---
title: "ROI Linear Mixed Models – ACC, Amy, dlPFC"
author: ""
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_float: true
    code_folding: show
    theme: darkly
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(file.path(dirname(rstudioapi::getSourceEditorContext()$path))))
```

# Overview

One **linear mixed model per region** (ACC, Amygdala, dlPFC) with hemisphere included as a within-subject factor instead of fitting 6 separate models.

**Model per region:** `mean_contrast ~ group * primer * target * hemisphere + (1 | subject)`

-   **Group** (between): HC vs MDD
-   **Primer** (within): happy vs sad
-   **Target** (within): happy vs sad
-   **Hemisphere** (within): L vs R

Type III F-tests via `lmerTest` (Satterthwaite ddf).
Bonferroni correction across 3 regions.

# Load data & packages

```{r load}
library(lme4)
library(lmerTest)
library(emmeans)
library(dplyr)
library(readr)
library(knitr)

out_dir <- "/Users/julia/Library/CloudStorage/OneDrive-UniklinikRWTHAachen/Paper/VideoMask/05_Cortex/revision1/new_mri/ROI/results"
csv_path <- "/Users/julia/Library/CloudStorage/OneDrive-UniklinikRWTHAachen/Paper/VideoMask/05_Cortex/revision1/new_mri/ROI/results/roi_condition_values_long.csv"
df <- read_csv(csv_path, show_col_types = FALSE)
df <- df[!is.na(df$mean_contrast), ]

# Factor coding
df$group      <- factor(df$group, levels = c("HC", "MDD"))
df$primer     <- factor(df$primer, levels = c("happy", "sad"))
df$target     <- factor(df$target, levels = c("happy", "sad"))
df$hemisphere <- factor(df$hemisphere, levels = c("L", "R"))
df$subject    <- factor(df$subject)
df$congruency <- factor(df$congruency, levels = c("congruent", "incongruent"))

# Extract region from ROI name (strip L_/R_ prefix)
df$region <- sub("^[LR]_", "", df$roi)

regions   <- sort(unique(df$region))
n_regions <- length(regions)

cat(sprintf("Loaded %d rows | %d subjects (HC=%d, MDD=%d) | Regions: %s\n",
            nrow(df), n_distinct(df$subject),
            n_distinct(df$subject[df$group == "HC"]),
            n_distinct(df$subject[df$group == "MDD"]),
            paste(regions, collapse = ", ")))
```

# 1 — Full Model {.tabset}

One LMM per region: `mean_contrast ~ group * primer * target * hemisphere + (1 | subject)`

```         
Bonferroni correction: p × r n_regions (regions).
```

```{r fit-models, results='asis'}
model_list    <- list()
anova_results <- list()

for (reg in regions) {
  d <- df[df$region == reg, ]

  mod <- lmer(mean_contrast ~ group * primer * target + hemisphere + (1 | subject),
              data = d)
  model_list[[reg]] <- mod

  aov <- anova(mod, type = 3)
  aov_df <- as.data.frame(aov)
  aov_df$Effect <- rownames(aov_df)

  p_colname <- grep("^Pr|^p", names(aov_df), value = TRUE)[1]
  aov_df$p_bonf <- pmin(aov_df[[p_colname]] * n_regions, 1.0)
  aov_df$sig <- ifelse(aov_df$p_bonf < 0.001, "***",
                ifelse(aov_df$p_bonf < 0.01,  "**",
                ifelse(aov_df$p_bonf < 0.05,  "*", "")))
  aov_df$Region <- reg
  anova_results[[reg]] <- aov_df

  cat(sprintf("\n\n## %s\n\n", reg))
  show_cols <- c("Effect", "Sum Sq", "Mean Sq", "NumDF", "DenDF", "F value",
                 p_colname, "p_bonf", "sig")
  show_cols <- show_cols[show_cols %in% names(aov_df)]
  print(kable(aov_df[, show_cols], digits = 4, row.names = FALSE,
              caption = sprintf("%s – Type III ANOVA (Bonferroni × %d)", reg, n_regions)))
}
```

## Significant effects (p_bonf \< .05)

```{r sig-effects}
anova_all <- bind_rows(anova_results)
p_colname <- grep("^Pr|^p", names(anova_all), value = TRUE)[1]
sig_effects <- anova_all[anova_all$p_bonf < 0.05, ]

if (nrow(sig_effects) > 0) {
  kable(sig_effects[, c("Region", "Effect", "F value", p_colname, "p_bonf", "sig")],
        digits = 4, row.names = FALSE,
        caption = sprintf("Significant effects (p_bonf < .05) — %d / %d",
                          nrow(sig_effects), nrow(anova_all)))
} else {
  cat("No significant effects at p_bonf < .05.")
}
```

# 2 — Post-hoc: HC vs MDD per Condition

Group comparisons (emmeans) at each Primer × Target combination per region.
Bonferroni correction: p × (4 conditions × `r n_regions` regions) = p × `r 4 * n_regions`.

```{r posthoc-group, results='asis'}
posthoc_results <- list()

for (reg in regions) {
  emm <- emmeans(model_list[[reg]], ~ group | primer * target)
  con <- pairs(emm, adjust = "none")
  con_df <- as.data.frame(con)
  con_df$Region <- reg

  n_posthoc <- n_regions * nrow(con_df)
  con_df$p_bonf <- pmin(con_df$p.value * n_posthoc, 1.0)
  con_df$sig <- ifelse(con_df$p_bonf < 0.001, "***",
                ifelse(con_df$p_bonf < 0.01,  "**",
                ifelse(con_df$p_bonf < 0.05,  "*", "")))
  posthoc_results[[reg]] <- con_df

  cat(sprintf("\n\n## %s\n\n", reg))
  print(kable(con_df[, c("Region", "primer", "target", "contrast", "estimate",
                          "SE", "t.ratio", "p.value", "p_bonf", "sig")],
              digits = 4, row.names = FALSE))
}
```

# 3 — Post-hoc: Hemisphere per Group

L vs R comparisons per group per region.
Bonferroni correction: p × (2 groups × `r n_regions` regions) = p × `r 2 * n_regions`.

```{r posthoc-hemi, results='asis'}
hemi_results <- list()

for (reg in regions) {
  emm <- emmeans(model_list[[reg]], ~ hemisphere | group)
  con <- pairs(emm, adjust = "none")
  con_df <- as.data.frame(con)
  con_df$Region <- reg

  n_hemi <- n_regions * nrow(con_df)
  con_df$p_bonf <- pmin(con_df$p.value * n_hemi, 1.0)
  con_df$sig <- ifelse(con_df$p_bonf < 0.001, "***",
                ifelse(con_df$p_bonf < 0.01,  "**",
                ifelse(con_df$p_bonf < 0.05,  "*", "")))
  hemi_results[[reg]] <- con_df

  cat(sprintf("\n\n## %s\n\n", reg))
  print(kable(con_df[, c("Region", "group", "contrast", "estimate",
                          "SE", "t.ratio", "p.value", "p_bonf", "sig")],
              digits = 4, row.names = FALSE))
}
```

# 4 — Congruency Model {.tabset}

`mean_contrast ~ group * congruency * hemisphere + (1 | subject)`

Averaged over the two congruent (HH, SS) and two incongruent (HS, SH) conditions per subject.
Bonferroni correction: p × `r n_regions` (regions).

```{r congruency, results='asis'}
df_cong <- df %>%
  group_by(subject, group, region, hemisphere, congruency) %>%
  summarise(mean_contrast = mean(mean_contrast), .groups = "drop")

cong_results <- list()

for (reg in regions) {
  d <- df_cong[df_cong$region == reg, ]
  mod <- lmer(mean_contrast ~ group * congruency + hemisphere + (1 | subject),
              data = d)

  aov <- anova(mod, type = 3)
  aov_df <- as.data.frame(aov)
  aov_df$Effect <- rownames(aov_df)
  p_colname <- grep("^Pr|^p", names(aov_df), value = TRUE)[1]
  aov_df$p_bonf <- pmin(aov_df[[p_colname]] * n_regions, 1.0)
  aov_df$sig <- ifelse(aov_df$p_bonf < 0.001, "***",
                ifelse(aov_df$p_bonf < 0.01,  "**",
                ifelse(aov_df$p_bonf < 0.05,  "*", "")))
  aov_df$Region <- reg
  cong_results[[reg]] <- aov_df

  cat(sprintf("\n\n## %s\n\n", reg))
  show_cols <- c("Effect", "F value", p_colname, "p_bonf", "sig")
  show_cols <- show_cols[show_cols %in% names(aov_df)]
  print(kable(aov_df[, show_cols], digits = 4, row.names = FALSE,
              caption = sprintf("%s – Congruency ANOVA (Bonferroni × %d)", reg, n_regions)))
}
```

# 5 — Export

```{r export}

write_csv(bind_rows(anova_results),   file.path(out_dir, "roi_lmm_anova.csv"))
write_csv(bind_rows(posthoc_results), file.path(out_dir, "roi_lmm_posthoc_group.csv"))
write_csv(bind_rows(hemi_results),    file.path(out_dir, "roi_lmm_posthoc_hemisphere.csv"))
write_csv(bind_rows(cong_results),    file.path(out_dir, "roi_lmm_congruency.csv"))

cat("Saved to results/:\n")
cat("  roi_lmm_anova.csv\n")
cat("  roi_lmm_posthoc_group.csv\n")
cat("  roi_lmm_posthoc_hemisphere.csv\n")
cat("  roi_lmm_congruency.csv\n")
```
